---
title: "Stepwise Salary Visualizations"
author: "4ward Acquisition"
date: "2026-04-16"
output:
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    theme: bootstrap
    css: dashboard_style.css
    navbar:
      - { title: "Josh Collins", align: right }
      - { title: "April 2026",   align: right }
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(tidyverse)
library(olsrr)
library(faraway)
library(leaps)
library(GGally)
library(glmnet)
library(quantreg)
library(kableExtra)
library(modelr)
library(htmltools)
library(scales)

knitr::opts_chunk$set(
  echo    = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2
)

dark_theme <- theme_minimal(base_size = 13) +
  theme(
    plot.background   = element_rect(fill = "#161b22", color = NA),
    panel.background  = element_rect(fill = "#161b22", color = NA),
    panel.grid.major  = element_line(color = "#1c2330"),
    panel.grid.minor  = element_blank(),
    plot.title        = element_text(color = "#e6edf3", face = "bold", size = 14),
    plot.subtitle     = element_text(color = "#8b949e", size = 11),
    axis.text         = element_text(color = "#8b949e"),
    axis.title        = element_text(color = "#8b949e"),
    legend.background = element_rect(fill = "#161b22", color = NA),
    legend.text       = element_text(color = "#8b949e"),
    legend.title      = element_text(color = "#8b949e"),
    strip.text        = element_text(color = "#8b949e"),
    plot.margin       = margin(16, 20, 16, 20)
  )
```

```{r data-prep}
combined <- read_csv("data/clean/combined.csv")

combined_clean <- combined %>%
  filter(!is.na(SALARY), !is.na(AGY), !is.na(EDLVL), !is.na(LOS)) %>%
  droplevels()

combined_clean$SALARY <- combined_clean$SALARY / 1000

model_df <- combined_clean %>%
  select(-GSEGRD, -WORKSCH, -WORKSTAT, -PP, -PPGRD, -TOA, -SALLVLT, -AGYSUBT, -AGYSUB) %>%
  mutate(
    LOC    = ifelse(grepl("^[0-9]+$", LOC), LOC, "57"),
    LOC    = as.numeric(LOC),
    EDLVL  = na_if(EDLVL, "**"),
    EDLVL  = as.numeric(EDLVL),
    LOSLVL = na_if(LOSLVL, "Z"),
    LOSLVL = match(LOSLVL, c("A","B","C","D","E","F","G","H","I","J")),
    AGELVL = match(AGELVL, c("B","C","D","E","F","G","H","I","J","K"))
  ) %>%
  na.omit() %>%
  droplevels()

bad_vars <- names(which(sapply(model_df, function(x) length(unique(x)) < 2)))
model_df <- model_df %>% select(-all_of(bad_vars)) %>% droplevels()

null_model   <- lm(SALARY ~ 1, data = model_df)
upper_model  <- lm(SALARY ~ ., data = model_df)
n            <- nrow(model_df)

forward_bic  <- step(null_model, direction = "forward",
                     scope = formula(upper_model), k = log(n), trace = 0)
step_bic     <- step(null_model, direction = "both",
                     scope = formula(upper_model), k = log(n), trace = 0)
simple_model <- lm(SALARY ~ LOS + AGELVL + EDLVL, data = model_df)

plot_df <- model_df %>%
  add_predictions(step_bic, var = "pred") %>%
  add_residuals(step_bic,   var = "resid")

agency_levels <- levels(factor(plot_df$AGY))
pal <- colorRampPalette(
  c("#58a6ff","#3fb950","#f78166","#d2a8ff","#e3b341",
    "#79c0ff","#56d364","#ffa657","#ff7b72","#b392f0")
)(length(agency_levels))
names(pal) <- agency_levels
```


```{r}
com <- combined_clean %>%
  lm(SALARY ~ LOS + + SUPERVIS + DATECODE + EDLVL + 
    AGY + AGELVL + LOC)
```

<!-- ############################################################
     PAGE 1 — MODEL SUMMARY  (columns layout: stats pinned left)
     ############################################################ -->

Model Summary {data-orientation=columns}
=====================================

Column {data-width=340 .sidebar}
-------------------------------------

### Model at a Glance

```{r}
terms <- attr(terms(step_bic), "term.labels")
r2    <- round(summary(step_bic)$r.squared, 3)

htmltools::HTML(paste0('
  <div class="sidebar-stats">

    <div class="stat-block">
      <div class="stat-label">Observations</div>
      <div class="stat-value accent-blue">', scales::comma(n), '</div>
    </div>

    <div class="stat-block">
      <div class="stat-label">Selected Model</div>
      <div class="stat-value" style="font-size:1.1rem">Bidirectional BIC</div>
    </div>

    <div class="stat-block">
      <div class="stat-label">R-squared</div>
      <div class="stat-value accent-green">', r2, '</div>
      <div class="stat-sub">Model explains ', round(r2*100), '% of salary variance</div>
    </div>

    <div class="stat-block">
      <div class="stat-label">BIC</div>
      <div class="stat-value accent-purple">', scales::comma(round(BIC(step_bic),0)), '</div>
    </div>

    <div class="stat-block">
      <div class="stat-label">AIC</div>
      <div class="stat-value accent-gold">', scales::comma(round(AIC(step_bic),0)), '</div>
    </div>

    <div class="stat-divider"></div>

    <div class="stat-label">Predictors (', length(terms), ')</div>
    <div class="formula-terms" style="margin-top:.5rem">',
      paste(sprintf('<span class="term-badge">%s</span>', terms), collapse=" "),
    '</div>

    <div class="stat-divider"></div>

    <div class="stat-label" style="margin-bottom:.4rem">Selection method</div>
    <div class="formula-meta" style="flex-direction:column;gap:.35rem;border:none;padding:0">
      <span>Direction: <strong>both</strong></span>
      <span>Penalty: <strong>k = log(n)</strong></span>
      <span>Response: <strong>SALARY / 1000</strong></span>
    </div>

  </div>
'))
```

Column {data-width=860}
-------------------------------------

### Model Comparison

```{r}
comparison_tbl <- tibble(
  Model     = c("Forward BIC", "Bidirectional BIC", "Simple (LOS+AGE+ED)"),
  Variables = c(length(attr(terms(forward_bic), "term.labels")),
                length(attr(terms(step_bic),    "term.labels")),
                3L),
  BIC = c(BIC(forward_bic), BIC(step_bic), BIC(simple_model)),
  AIC = c(AIC(forward_bic), AIC(step_bic), AIC(simple_model)),
  R2  = c(summary(forward_bic)$r.squared,
          summary(step_bic)$r.squared,
          summary(simple_model)$r.squared)
)
best_row <- which.min(comparison_tbl$BIC)

comparison_tbl %>%
  mutate(BIC = scales::comma(round(BIC, 1)),
         AIC = scales::comma(round(AIC, 1)),
         R2  = round(R2, 4)) %>%
  kbl(align = c("l","c","r","r","r"), escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = TRUE, font_size = 14) %>%
  row_spec(0,        bold = TRUE, color = "#8b949e", background = "#161b22") %>%
  row_spec(best_row, bold = TRUE, color = "#3fb950", background = "#0f2318") %>%
  column_spec(1, bold = TRUE)
```

### About This Model

```{r}
htmltools::HTML('
  <div class="about-card">
    <div class="about-col">
      <div class="about-label">What was done</div>
      <p>Three linear regression models were estimated on federal employment salary data.
      Variable selection used stepwise regression penalised by BIC (<code>k = log(n)</code>),
      which applies a stricter parsimony penalty than AIC, especially at large sample sizes.
      The bidirectional search both adds and removes variables at each step.</p>
    </div>
    <div class="about-col">
      <div class="about-label">Why BIC over AIC</div>
      <p>With over 230,000 observations, AIC tends to retain more variables than necessary.
      BIC scaling by <code>log(n)</code> makes each additional predictor substantially more
      costly, producing a leaner model less likely to overfit noise in the training data.</p>
    </div>
    <div class="about-col">
      <div class="about-label">Interpreting R²</div>
      <p>The selected model explains approximately <strong>56%</strong> of variance in salary.
      The remaining 44% reflects individual negotiation, within-grade pay steps, performance
      bonuses, and other factors not present in the dataset.</p>
    </div>
  </div>
')
```


<!-- ############################################################
     PAGE 2 — RESIDUAL ANALYSIS  (rows layout: formula top, plots below)
     ############################################################ -->

Residual Analysis {data-orientation=rows}
=====================================

Row {data-height=90}
-------------------------------------

### {data-width=1200}
```{r}
htmltools::HTML('
  <div class="page-intro">
    <strong>Residual Analysis</strong> — Residuals (Actual minus Predicted) reveal where the
    model succeeds and where it struggles. Ideally residuals are centred at zero with symmetric
    spread and no systematic patterns across agencies or salary levels.
  </div>
')
```

Row {data-height=460}
-------------------------------------

### Residual Distribution {data-width=500}
```{r fig.height=4.2, fig.width=6}
plot_df %>%
  ggplot(aes(x = resid, fill = after_stat(x > 0))) +
  geom_histogram(bins = 60, color = "transparent") +
  scale_fill_manual(values = c("#58a6ff99","#f7816699"), guide = "none") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#e3b341", linewidth = .8) +
  labs(title    = "Distribution of Residuals",
       subtitle = "Blue = over-predicted  |  Orange = under-predicted",
       x = "Residual ($k)", y = "Count") +
  dark_theme
```

### Predicted vs. Actual {data-width=700}
```{r fig.height=4.2, fig.width=8}
plot_df %>%
  slice_sample(n = min(3000, nrow(plot_df))) %>%
  ggplot(aes(x = pred, y = SALARY, color = AGY)) +
  geom_point(alpha = 0.35, size = 1) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", color = "#e3b341", linewidth = .8) +
  scale_color_manual(values = pal) +
  labs(title    = "Predicted vs. Actual Salary",
       subtitle = "3,000 point sample  |  Dashed = perfect prediction",
       x = "Predicted ($k)", y = "Actual ($k)", color = "Agency") +
  dark_theme
```

Row {data-height=130}
-------------------------------------

### Residual Distribution — Interpretation {data-width=500}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">📊</div>
    <div><p>Residuals are roughly centred at zero as expected from OLS. The mild
    <strong>right skew</strong> is typical for salary data — a small number of very high
    earners are harder to predict, pulling the distribution rightward. A log transformation
    of SALARY may reduce this skew in future iterations.</p></div>
  </div>
')
```

### Predicted vs. Actual — Interpretation {data-width=700}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">🎯</div>
    <div><p>Points cluster along the 45° line confirming the model captures the central salary
    trend. The <strong>spread widens at higher salaries</strong> — a classic sign of
    heteroscedasticity. Senior and specialised roles have more individual variance not
    explained by the administrative predictors.</p></div>
  </div>
')
```

Row {data-height=460}
-------------------------------------

### Residuals by Agency {data-width=1200}
```{r fig.height=4.2, fig.width=13}
plot_df %>%
  group_by(AGY) %>%
  slice_sample(n = 500, replace = TRUE) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(AGY, resid, median), y = resid,
             fill = AGY, color = AGY)) +
  geom_violin(alpha = 0.5, linewidth = .4) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "#e3b341", linewidth = .7) +
  scale_fill_manual(values  = pal, guide = "none") +
  scale_color_manual(values = pal, guide = "none") +
  labs(title    = "Residual Distribution by Agency",
       subtitle = "Ordered by median residual  |  Dashed = zero  |  500 obs/agency sample",
       x = "Agency", y = "Residual ($k)") +
  dark_theme
```

Row {data-height=130}
-------------------------------------

### Residuals by Agency — Interpretation {data-width=1200}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">🏛️</div>
    <div><p>Agencies whose violins sit <strong>above zero</strong> are systematically
    under-predicted. Those <strong>below zero</strong> are over-predicted. Wide violins
    indicate high within-agency variance in prediction error, which may reflect heterogeneous
    job types or pay scales not fully captured by the current predictors.</p></div>
  </div>
')
```


<!-- ############################################################
     PAGE 3 — SALARY PATTERNS  (rows layout: formula top, plots below)
     ############################################################ -->

Salary Patterns {data-orientation=rows}
=====================================

Row {data-height=90}
-------------------------------------

### {data-width=1200}
```{r}
htmltools::HTML('
  <div class="page-intro">
    <strong>Salary Patterns</strong> — Exploring how salary varies with length of service,
    agency, and education level, and how well the model captures those patterns.
  </div>
')
```

Row {data-height=460}
-------------------------------------

### Salary vs. Length of Service {data-width=1200}
```{r fig.height=4.2, fig.width=13}
model_df %>%
  slice_sample(prop = 0.08) %>%
  ggplot(aes(x = LOS, y = SALARY, color = AGY)) +
  geom_point(alpha = 0.08, size = .8) +
  geom_smooth(data = model_df, method = "lm", se = FALSE, linewidth = 1.2) +
  scale_color_manual(values = pal) +
  labs(title    = "Salary vs. Length of Service by Agency",
       subtitle = "Regression line per agency  |  Points = 8% random sample",
       x = "Length of Service (years)", y = "Salary ($k)", color = "Agency") +
  dark_theme
```

Row {data-height=130}
-------------------------------------

### Length of Service — Interpretation {data-width=1200}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">📈</div>
    <div><p>Each line shows the agency-specific salary growth rate with seniority.
    <strong>Steeper slopes</strong> indicate a stronger seniority pay premium within that agency.
    The diverging slopes across agencies supports including both <code>LOS</code> and
    <code>AGY</code> as separate predictors. An interaction term <code>LOS:AGY</code>
    would allow the model to capture these differing slopes directly.</p></div>
  </div>
')
```

Row {data-height=460}
-------------------------------------

### Predicted vs. Actual by Agency {data-width=700}
```{r fig.height=4.2, fig.width=8}
plot_df %>%
  group_by(AGY) %>%
  slice_sample(n = 300, replace = TRUE) %>%
  ungroup() %>%
  select(AGY, SALARY, pred) %>%
  pivot_longer(c(SALARY, pred), names_to = "Type", values_to = "Salary") %>%
  mutate(Type = recode(Type, SALARY = "Actual", pred = "Predicted")) %>%
  ggplot(aes(x = AGY, y = Salary, fill = Type)) +
  geom_violin(alpha = 0.6, position = position_dodge(.9), linewidth = .3) +
  scale_fill_manual(values = c(Actual = "#58a6ff99", Predicted = "#3fb95099")) +
  labs(title    = "Predicted vs. Actual Salary by Agency",
       subtitle = "Blue = Actual  |  Green = Predicted  |  300 obs/agency sample",
       x = "Agency", y = "Salary ($k)", fill = NULL) +
  dark_theme
```

### Salary by Education Level {data-width=500}
```{r fig.height=4.2, fig.width=6}
model_df %>%
  group_by(EDLVL) %>%
  summarise(avg_salary = mean(SALARY, na.rm = TRUE), n = n(), .groups = "drop") %>%
  arrange(EDLVL) %>%
  ggplot(aes(x = factor(EDLVL), y = avg_salary)) +
  geom_col(fill = "#d2a8ff55", color = "#d2a8ff", linewidth = .6) +
  geom_line(aes(group = 1), color = "#d2a8ff", linewidth = 1.1) +
  geom_point(color = "#d2a8ff", size = 3) +
  labs(title    = "Average Salary by Education Level",
       subtitle = "Higher codes = more advanced qualifications",
       x = "EDLVL Code", y = "Average Salary ($k)") +
  dark_theme
```

Row {data-height=160}
-------------------------------------

### Predicted vs. Actual — Interpretation {data-width=700}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">🎻</div>
    <div><p>Paired violins compare actual (blue) and predicted (green) salary distributions
    per agency. A <strong>narrower predicted violin</strong> indicates the model compresses
    extreme values — capturing the centre but missing the tails, a known limitation of OLS
    minimising squared error.</p></div>
  </div>
')
```

### Education Level — Interpretation {data-width=500}
```{r}
htmltools::HTML('
  <div class="interp-card">
    <div class="interp-icon">🎓</div>
    <div><p>The monotonically increasing trend confirms education is a meaningful positive
    predictor of federal salary. The <strong>steeper rise at higher EDLVL codes</strong>
    suggests non-linear returns to education. A quadratic term or factor encoding of EDLVL
    may improve fit beyond the current linear treatment.</p></div>
  </div>
')
```
